Type-Safe Client-Server Communication in Web Applications

### 2014-02-27 paper.tex
Add half of the Haste.App chatbox example to paper.tex.

Now that we have seen how to implement both network communication, we are ready
to revisit the chatbox program, this time using our improved programming model.

>>>>

Now that we have seen how to implement both network communication, we are ready
to revisit the chatbox program, this time using our improved programming model.
Since we are now writing the entire application, both client and server, as
opposed to the client part from our motivating example, our program has three
new responsibilities.

\begin{itemize}
  \item We need to add connecting users to a list of message recipients;
  \item users leaving the site need to be removed from the recipient list; and
  \item chat messages need to be distributed to all users in the list.
\end{itemize}

With this in mind, we begin by importing a few modules we are going to need and
define the type for our recipient list.

\begin{code}
import Data.List (find)
import Haste.App
import Haste.App.Concurrent
import qualified Control.Concurrent as CC

type Recipient = (SessionID, CC.MVar String)
type RcptList = CC.MVar [Recipient]
\end{code}

We use an \lstinline!MVar! from \lstinline!Control.Concurrent! to store the
list of recipients. A recipient will be represented by a \lstinline!SessionID!,
an ID used by Haste.App to identify user sessions, and an \lstinline!MVar! into
which new chat messages sent to the recipient will be written. Next, we define
our handshake RPC function.

\begin{code}
srvHello :: Useless RcptList -> Server ()
srvHello uRcpts = do
  recipients <- mkUseful uRcpts
  sid <- getSessionID
  rcptMVar <- liftIO CC.newEmptyMVar
  liftIO $ CC.modifyMVar recipients $ \cs -> do
    return ((sid, rcptMVar):cs ,())
\end{code}

The handshake is a simple operation; the connecting client just gets an
\lstinline!MVar! associated with its session identifier, and the two are then
prepended to the recipient list. Notice how the application's server state is
passed in as the function's first \lstinline!Useless RcptList! argument.

\begin{code}
srvSend :: Useless RcptList -> String -> Server ()
srvSend uRcpts message = do
    rcpts <- mkUseful uRcpts
    liftIO $ do
      recipients <- CC.withMVar rcpts return
      mapM_ (CC.forkIO . deliver message) recipients
  where
    deliver :: String -> Recipient -> IO ()
    deliver message (_, rcptMVar) =
      CC.putMVar rcptMVar message
\end{code}

The send function is slightly more complex. The incoming message is written to
the \lstinline!MVar! corresponding to each active session. Note the use of
\lstinline!forkIO!. Some recipients may be slow to empty their \lstinline!MVar!
to make room for a new message. We don't want to force all recipients to wait
for the slowest one, so we spawn a new lightweight thread for each recipient.

\begin{code}
srvAwait :: Useless RcptList -> Server String
srvAwait uRcpts = do
  rcpts <- mkUseful uRcpts
  sid <- getSessionID
  liftIO $ do
    recipients <- CC.withMVar rcpts return
    case find ((== sid) . fst) recipients of
      Just (_, mv) -> CC.takeMVar mv
      _            -> fail "Unregistered session; aborting"
\end{code}

The final server operation, notifying users of pending messages, finds the
appropriate \lstinline!MVar! to wait on by searching the recipient list for the
session identifier of the calling user, and then blocks until a message arrives
in said \lstinline!MVar!. This is a little different from the other two
operations, which perform their work as quickly as possible and then return
immediately.
